### 2013-10-28 Talks/FU-Berlin/GödelGod-Talk.tex
FU talk

\documentclass[9pt]{beamer}
\mode<presentation>
{
  \usetheme{Warsaw}
  \usecolortheme{crane}
  \setbeamercovered{transparent}
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{verbatim}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{pxfonts}
\usepackage{amsthm}
\usepackage{graphicx}

\usepackage{bussproofs}
\usepackage{proof}

\usepackage{fancybox}
\usepackage{fancyvrb}

%\usepackage{rotating}

% Sequent Calculus Proof Settings
\EnableBpAbbreviations
\def\fCenter{\mbox{\ $\vdash$\ }}

\newcommand{\mypause}{\pause}
%\newcommand{\mypause}{}

\title[On G\"{o}del's Proof of God's
Existence]{Formalization, Mechanization and Automation of \\ G\"{o}del's
  Proof of God's Existence}

\author{Christoph Benzm\"{u}ller and Bruno Woltzenlogel Paleo}

\institute[]{
  \inst{}%
}

\date[1.11.2013]
{November 1, 2013}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\input{talk-part-1.tex}

\input{talk-part-2.tex}

\input{talk-part-3.tex}

\input{talk-part-4.tex}

\input{talk-part-5.tex}

\input{talk-part-6.tex}

\end{document}