### 2013-10-29 Talks/FU-Berlin/talk-part-4.tex

\begin{frame}{Coq Proof} \small
todo
\end{frame}