### 2013-10-29 Talks/FU-Berlin/talk-part-4.tex
Add Coq slide

\begin{frame}{Coq Proof}{Demo} \small
\begin{itemize}
\item Goal: verification of the natural deduction proof
\begin{itemize}
\item Step-by-step formalization
\item Almost no automation (intentionally!)
\end{itemize}
%
\item Interesting facts to note:
\begin{itemize}
\item Embedding is transparent to the user
\item Embedding gives labeled calculus for free
\end{itemize}
\end{itemize}
\end{frame}