### 2013-10-29 Talks/FU-Berlin/GödelGod-Talk.tex

\documentclass[9pt]{beamer}
\mode<presentation>
{
  \usetheme{Warsaw}
  \usecolortheme{crane}
  \setbeamercovered{transparent}
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{verbatim}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{pxfonts}
\usepackage{amsthm}
\usepackage{graphicx}

\usepackage{bussproofs}
\usepackage{proof}

\usepackage{fancybox}
\usepackage{fancyvrb}

%\usepackage{rotating}

% Sequent Calculus Proof Settings
\EnableBpAbbreviations
\def\fCenter{\mbox{\ $\vdash$\ }}

\usepackage{calculi}

% Logical symbols
\newcommand{\imp}{\rightarrow}
\newcommand{\biimp}{\leftrightarrow}
\newcommand{\all}{\forall}
\newcommand{\ex}{\exists}
\newcommand{\seq}{\vdash}
\newcommand{\nec}{\Box} % necessarily
\newcommand{\pos}{\Diamond} % possibly

\newcommand{\ess}[2]{#1 \ \mathit{ess} \ #2}
\newcommand{\NE}{E}

\newcommand{\s}{\qquad}

\newcommand{\mypause}{\pause}
%\newcommand{\mypause}{}

\title[On G\"{o}del's Proof of God's
Existence]{Formalization, Mechanization and Automation of \\ G\"{o}del's
  Proof of God's Existence}

\author{\textbf{Christoph Benzm\"{u}ller} and \textbf{Bruno Woltzenlogel Paleo}}

% \institute[]{
%   \inst{}%
% }

\date[1.11.2013]{November 1, 2013}

\begin{document}

\begin{frame}
  \titlepage
\includegraphics[height=2.5cm]{buch7.jpg} 
\hfill
\includegraphics[height=2.5cm]{nd.jpg}
\end{frame}

\input{talk-part-1.tex}

\input{talk-part-2.tex}

\input{talk-part-3.tex}

\input{talk-part-4.tex}

\input{talk-part-5.tex}

\input{talk-part-6.tex}

\end{document}