### 2013-10-30 Talks/FU-Berlin/talk-part-6.tex

\begin{frame}{Criticisms}{S5}

\end{frame}

\begin{frame}{Criticisms}{Modal Collapse}

\end{frame}

\begin{frame}{Criticisms}{No Neutral Properties}

\end{frame}

\begin{frame}{Summary of Results}
\begin{itemize}[<+->]
\item K sufficient for T1, C and T2 
\item S5 not needed for T3
\item KB sufficient for T3 
\item G\"odel's original axioms (without conjunct $\phi(x)$ in D2) were inconsistent
\item Scott's axioms are consistent
\item For T1, only half of A1 (A1a) is needed 
\item For T2, the other half (A1b) is needed
\end{itemize}
\end{frame}

\begin{frame}{Conclusion} \small
todo
\end{frame}

\begin{frame}[plain]
\includegraphics[width=\textwidth]{TUWien-GodComputerC} 
\end{frame}