### 2013-10-30 Talks/FU-Berlin/talk-part-5.tex

\begin{frame}{} \Large \centering
\includegraphics[width=.7\textwidth]{TUWien-GodComputerC} 
\vfill
\textbf{Part D:} \\
\quad automation \& verification: proof assistant \textsc{Isabelle} \\[2em]
\end{frame}

\begin{frame}{} \Large \centering
\vskip.5em
\colorbox{gray}{\includegraphics[width=\textwidth]{IsabelleGrab}} 
\end{frame}

\begin{frame}{Automation \& Verification in Proof Assistant \textsc{Isabelle/HOL}} \Large
Isabelle
\begin{itemize}
\item bla
\end{itemize}
\vfill
What have we done
\begin{itemize}
\item bla 
\end{itemize}
\vfill
See the handout (generated from the Isabelle source file).
\end{frame}