### 2013-10-31 Talks/FU-Berlin/talk-part-5.tex
update of FU talk

\begin{transitionframe}{} \Large \centering
\textbf{Part D:} \\
\quad automation \& verification: proof assistant \textsc{Isabelle} \\[2em]
\end{transitionframe}

\begin{frame}{} \Large \centering
\vskip.5em
\colorbox{gray}{\includegraphics[width=\textwidth]{IsabelleGrab}} 
\end{frame}

\begin{frame}{Automation \& Verification in Proof Assistant \textsc{Isabelle/HOL}} \Large
Isabelle/HOL   (Cambridge University/TU Munich)
\begin{itemize}
\item HOL instance of the generic \textsc{Isabelle} proof assistant
\item User interaction and proof automation 
\item Automation is supported by \textsc{Sledgehammer} tool
\item Verification of the proofs in \textsc{Isabelle/HOL}'s small proof kernel
\end{itemize}
\vfill
What have we done?
\begin{itemize}
\item Proof automation of G\"odel's proof script (Scott version)
\item \textsc{Sledghammer} makes calls to remote THF provers in Miami
\item These calls the suggest respective calls to the \textsc{Metis} prover
\item \textsc{Metis} proofs are verified in \textsc{Isabelle/HOL}'s proof kernel
\end{itemize}
\vfill
See the handout (generated from the Isabelle source file).
\end{frame}