### 2013-10-31 Talks/FU-Berlin/talk-part-7.tex
update for FU talk

\begin{transitionframe}
\textbf{Part F:}

Conclusions
\end{transitionframe}

\begin{frame}{Summary of Results} \large

The (\alert{new}) insights we gained from experiments include:\\[.5em]
\begin{itemize}
\item Logic K sufficient for T1, C and T2 
\item Logic S5 not needed for T3
\item \alert{Logic KB sufficient for T3 (not well known)}
\item \alert{We found a simpler new proof of C}
\item \alert{G\"odel's axioms (without conjunct $\phi(x)$ in D2) are inconsistent}
\item Scott's axioms are consistent
\item For T1, only half of A1 (A1a) is needed 
\item For T2, the other half (A1b) is needed
\end{itemize}
\end{frame}

\begin{frame}{Summary of Results} \large

Our novel contributions the  theorem proving community includes \\[.5em]
\begin{itemize}
\item Powerful infrastructure for reasoning with higher-order modal logic using existing proof assistants and HOL provers
\item A new natural deduction calculus for higher-order modal logic
\item Difficult new benchmarks problems for HOL provers
\item Huge media attention
\end{itemize}
\end{frame}

\begin{frame}{Conclusion} \large
\vskip-1em What have we achieved \\[.5em]
\begin{itemize}
\item Verification of G\"odel's ontological argument with HOL provers
  \begin{itemize}
  \item exact parameters known: constant domain quantification, Henkin Semantics
  \item parameters can be varied and experiments can repeated
  \end{itemize}
\item Gained some novel results and insights
\item Major  step towards \alert{Computer-assisted Theoretical Philosophy}
 \begin{itemize}
  \item see also Ed Zalta's \emph{Computational Metaphysics} project at Stanford University
  \item remember Leibniz' dictum --- \emph{Calculemus!}
  \end{itemize}
\item Interesting bridge between CS, Philosophy and Theology
\end{itemize}

\pause

\vfill
\vskip-1em Ongoing and future work \\[.5em]
\begin{itemize}
\item Formalize and verify literature on ontological arguments
  \begin{itemize}
  \item \ldots in particular the criticism and improvements to G\"odel 
  \end{itemize}
\item Own contributions --- supported by theorem provers
\end{itemize}
\end{frame}

\begin{frame}{Some Comments and Reactions}
\colorbox{gray}{\includegraphics[width=.8\textwidth]{Comment1}}\\[.7em]

\, \hfill \colorbox{gray}{\includegraphics[width=.7\textwidth]{Comment2}}\\[.7em]

\colorbox{gray}{\includegraphics[width=.8\textwidth]{Comment3}}\\[1em]

\, \hfill \ldots find more on the internet \ldots
\end{frame}

\begin{frame}[plain]
\colorbox{black}{\includegraphics[width=\textwidth]{TUWien-GodComputerC}}
\end{frame}