### 2013-11-11 Formalizations/Isabelle/ScottsVersion/output/document/GoedelGodModalCollapse.tex
Improve Isabelle paper

%
\begin{isabellebody}%
\def\isabellecontext{GoedelGodModalCollapse}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isamarkupsection{Modal Collapse%
}
\isamarkuptrue%
%
\begin{isamarkuptext}%
G\"odel's axioms have been criticized for entailing the so-called 
modal collapse. The prover Satallax \cite{Satallax} confirms this. 
However, sledgehammer is not able to determine which axioms, 
definitions and previous theorems are used by Satallax;
hence it suggests to call Metis using everything, but this (unsurprinsingly) fails.
Attempting to use `Sledegehammer min' to minimize Sledgehammer's suggestion does not work.
Calling Metis with \isa{T{\isadigit{2}}}, \isa{T{\isadigit{3}}} and \isa{ess{\isacharunderscore}def} also does not work.%
\end{isamarkuptext}%
\isamarkuptrue%
\ \ \isacommand{lemma}\isamarkupfalse%
\ MC{\isacharcolon}\ {\isachardoublequoteopen}{\isacharbrackleft}{\isasymforall}{\isacharparenleft}{\isasymlambda}{\isasymPhi}{\isachardot}{\isacharparenleft}{\isasymPhi}\ m{\isasymrightarrow}\ {\isacharparenleft}{\isasymbox}\ {\isasymPhi}{\isacharparenright}{\isacharparenright}{\isacharparenright}{\isacharbrackright}{\isachardoublequoteclose}\isanewline
\ \ \isacommand{sledgehammer}\isamarkupfalse%
\ {\isacharbrackleft}provers\ {\isacharequal}\ remote{\isacharunderscore}satallax{\isacharbrackright}%
\isadelimproof
\ %
\endisadelimproof
%
\isatagproof
\isacommand{oops}\isamarkupfalse%
\isanewline
\ \ \isanewline
\isanewline
%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
\end{isabellebody}%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "root"
%%% End: