### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Write section 3 (tactics for modalities)

\documentclass{llncs}
\pagestyle{plain}

\usepackage{fancybox} % Must be before "fancyvrb"
\usepackage[utf8]{inputenc}
\usepackage{url,amsmath,amssymb,fancyvrb,color}
\usepackage{proof}
\usepackage{xspace}

\usepackage{calculi}

\newcommand{\imp}{\rightarrow}
\newcommand{\biimp}{\leftrightarrow}
\newcommand{\all}{\forall}
\newcommand{\ex}{\exists}
\newcommand{\seq}{\vdash}
\newcommand{\nec}{\Box} % necessarily
\newcommand{\pos}{\Diamond} % possibly

\newcommand{\red}[1]{\textcolor[rgb]{1,0,0}{#1}}
\newcommand{\blue}[1]{\textcolor[rgb]{0,0,1}{#1}}
\newcommand{\brown}[1]{\textcolor[rgb]{0.8,0.6,0.4}{#1}}

\newcommand{\Parameter}{\red{Parameter}}
\newcommand{\Ltac}{\red{Ltac}}
\newcommand{\Axiom}{\red{Axiom}}
\newcommand{\Lemma}{\red{Lemma}}
\newcommand{\Theorem}{\red{Theorem}}
\newcommand{\Definition}{\red{Definition}}
\newcommand{\Notation}{\blue{Notation}}
\newcommand{\Prop}{\blue{Prop}}
\newcommand{\Type}{\blue{Type}}
\newcommand{\llet}{\blue{let}}
\newcommand{\match}{\blue{match}}
\newcommand{\with}{\blue{with}}
\newcommand{\eend}{\blue{end}}
\newcommand{\iin}{\blue{in}}
\newcommand{\as}{\blue{as}}
\newcommand{\Require}{\blue{Require}}
\newcommand{\Import}{\blue{Import}}
\newcommand{\fforall}{\blue{forall}}
\newcommand{\fun}{\blue{fun}}
\newcommand{\Proof}{\blue{Proof}}
\newcommand{\Qed}{\blue{Qed}}
\newcommand{\Hint}{\blue{Hint}}
\newcommand{\com}[1]{\brown{#1}}
\newcommand{\bslash}{\symbol{92}}

\newcommand{\Coq}{\texttt{Coq}\xspace}
\newcommand{\Isabelle}{\texttt{Isabelle}\xspace}

\title{
Modal Logics in the Coq Proof Assistant
}

\author{
  Christoph Benzm\"{u}ller\inst{1}\thanks{This work has been supported by the German Research Foundation (DFG) under grant BE2501/9-1.} 
  \and 
  Bruno Woltzenlogel Paleo\inst{2}
}

\authorrunning{C.\~Benzm\"{u}ller \and B.\~Woltzenlogel Paleo}

\institute{
  Dahlem Center for Intelligent Systems, Freie Universit\"{a}t Berlin, Germany\\
  \email{c.benzmueller@gmail.com}
  \and 
  Theory and Logic Group, Vienna University of Technology, Austria \\
  \email{bruno@logic.at}
}

\begin{document}

\maketitle

\begin{abstract}  
This paper describes an embedding of higher-order
modal logics in the \Coq proof assistant. {\Coq}'s capabilities are
thus extended in a minimalistic manner, which is nevertheless
sufficient for the formalization of significant modal proofs. The
elegance, flexibility and convenience of this approach, from a user
perspective, are illustrated here with the successful formalization of
Gödel's ontological argument.  
\end{abstract}

\section{Introduction}

Modal logics extend the usual formal logic language with modal
operators ($\Box$ and $\Diamond$) and are characterized by the
\emph{necessitation rule}, according to which $\Box A$ is a theorem if
$A$ is a theorem, although $A \imp \Box A$ is not necessarily a
theorem. Various informal concepts, such as \emph{necessity and
possibility}, \emph{knowledge and belief}, \emph and \emph{temporal
globality and eventuality}, have been formalized with the help of
modal operators. Therefore, modal logics \cite{ModalLogic} are well-
established for a wide variety of application domains, ranging from
philosophy \cite{ToDo} to formal software verification
\cite{ToDo:TLAPlus}.

However, general automated reasoning support for modal logics is still
not as well-developed as for classical logics. Deduction tools
for modal logics are often limited to propositional, quantifier-free,
fragments \cite{ToDo:SeeBjornsProposal} or tailored to particular
modal logics and their applications \cite{ToDo:StephanMerz}; 
first-order automated deduction techniques based on tableaux, sequent
calculi and connection calculi have only recently been generalized 
and implemented in a few new provers able to directly cope with
modalities \cite{JensOtten}.

Another recently explored possibility \cite{J23,B9,J22} is the
embedding of first-order and even higher-order modal logics into classical
higher-order logics, for which existing higher-order automated theorem
provers \cite{ToDo:Leo,Satallax} exist.  Embedding is flexible,
because various modal logics can be easily supported by adding their
characteristic axioms.  The embedding approach can also be easily
adapted to support multiple modalities, and it is possible to replace
constant domain quantifiers by varying or cumulative domain
quantifiers.  Moreover, the approach is relatively simple to
implement, because it does not require any modification in the source
code of the higher-order prover. The prover can be used as is, and
only the input files provided to the prover must be especially
encoded. Furthermore, the efficacy and efficiency of the embedding
approach has been confirmed in philosophically interesting and
relevant benchmarks \cite{FormalTheologyRepository}. These qualities
make embedding a convenient approach for \emph{fully automated}
reasoning.

However, one may wonder whether the embedding approach is adequate
also for \emph{interactive} reasoning, when the user proves theorems
by interacting with a proof assistant such as \Coq\footnote{The \Coq
proof assistant was chosen merely because of the authors' greater
familiarity with the tactic language of this system. Nevertheless, the
techniques presented and discussed here are likely to be useful for
other proof assistants, such as \texttt{Isabelle} \cite{Isabelle} or
\texttt{HOL-Light} \cite{HOLLight}, as well.}. The main goal of this
paper is to study this question. Our answer is positive.

One major initial concern was whether the embedding could be a
disturbance to the user. Fortunately, by using \Coq's \texttt{Ltac}
tactic language, we were able to define intuitive new tactics that
hide the technical details of the embedding from the user. The
resulting infra-structure for modal reasoning within \Coq (as
described in Section \ref{sec:modal}) provides a user experience where
modalities can be handled transparently and straightforwardly.
Therefore, a user with basic knowledge of modal logics and \Coq's
tactics should be able to use (and extend) our implementation with no
excessive overhead.

In order to illustrate the use of the implemented embedding, we show
here the formalization of Scott's version \cite{Scott} of Gödel's
ontological argument for God's existence (in Section \ref{sec:proof}).
This proof was chosen mainly for two reasons. Firstly, it requires not
only modal operators, but also higher-order quantification. Therefore,
it is beyond the reach of specialized propositional and first-order
(modal) theorem provers. Secondly, this argument addresses an ancient
problem in Philosophy and Metaphysics, which has nevertheless received
a lot of attention in the last 15 years, because of the discovery of
the modal collapse \cite{ToDo}. This proof lies in the center of a
vast and largely unexplored application domain for automated and
interactive theorem provers.

The simpler ontological argument of Anselm has been automatically
verified with \texttt{PVS} by Rushby \cite{Rushby} and with first-
order theorem provers by Oppenheimer and Zalta \cite{Zalta}. Gödel's
argument was automatically verified in our previous work on embedding-
based fully automated modal theorem proving \cite{ArXiv,AFP}. But this
paper presents the first fully interactive and detailed formalization
of this proof in a proof assistant.

\section{The Embedding of Modal Logics in Coq}
\label{sec:modal}

A crucial aspect of modal logics \cite{ModalLogic} is that the so-
called \emph{necessitation rule} allows $\Box A$ to be derived if $A$
is a theorem, but $A \imp \Box A$ is not necessarily a theorem. Naive
attempts to define the modal operators $\Box$ and $\Diamond$ may
easily be unsound in this respect. To avoid this issue, the
\emph{possible world semantics} of modal logics can be explicitly
embedded into higher-order logics \cite{J23,B9}.

The embedding is related to labeling techniques \cite{Labels}.
However, the expressiveness of higher-order logic is exploited here to
encode the labels within the logical language. To this aim, a type for
worlds must be declared and modal propositions should be not of type
\texttt{Prop} but of a lifted type \texttt{o} that depends on possible
worlds:

\newcommand{\verbsize}{\small}

\begin{small}
\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Parameter i: \Type. \com{(* Type for worlds *)}
\Parameter u: \Type. \com{(* Type for individuals *)}
\Definition o := i -> \Prop. \com{(* Type of modal propositions *)}
\end{Verbatim}
\end{small}

\noindent
Possible worlds are connected by an accessibility relation, 
which can be represented in \Coq by a parameter \texttt{r}, as follows:

\begin{small}
\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Parameter r: i -> i -> \Prop. \com{(* Accessibility relation for worlds *)}
\end{Verbatim}
\end{small}

\noindent 
All modal connectives are simply lifted versions of the
usual logical connectives. Notations are used to allow the modal
connectives to be used as similarly as possible as the usual
connectives. The prefix ``\texttt{m}'' is used to distinguish the
modal connectives: if $\odot$ is a connective on type \text{Prop},
$\texttt{m}\odot$ is a connective on the lifted type $\texttt{o}$ of
modal propositions.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Definition mequal (x y: u)(w: i) := x = y.
\Notation "x m= y" := (mequal x y) (at level 99, right associativity).

\Definition mnot (p: o)(w: i) := ~ (p w).
\Notation "m~  p" := (mnot p) (at level 74, right associativity).

\Definition mand (p q:o)(w: i) := (p w) /\bslash (q w).
\Notation "p m/\bslash q" := (mand p q) (at level 79, right associativity).

\Definition mor (p q:o)(w: i) := (p w) \bslash\slash (q w).
\Notation "p m\bslash\slash q" := (mor p q) (at level 79, right associativity).

\Definition mimplies (p q:o)(w:i) := (p w) -> (q w).
\Notation "p m-> q" := (mimplies p q) (at level 99, right associativity).

\Definition mequiv (p q:o)(w:i) := (p w) <-> (q w).
\Notation "p m<-> q" := (mequiv p q) (at level 99, right associativity).
\end{Verbatim}

\noindent 
Likewise, modal quantifiers are lifted versions of the usual
quantifiers. \Coq's type system with dependent types is particularly
helpful here. The modal quantifiers \texttt{A} and \texttt{E} are
defined as depending on a type \texttt{t}. Therefore, they can
quantify over variables of any type. Moreover, the curly brackets
indicate that \texttt{t} is an implicit argument that can be inferred
by \Coq's type inference mechanism. This allows notations\footnote{The
keyword \texttt{\fun} indicates a lambda abstraction: \texttt{\fun \ x
=> p} (or \texttt{\fun \ x:t => p}) denotes the function 
$\lambda x:t.p$, which takes an argument $x$ (of type $t$) and returns $p$.}
(i.e. \texttt{mforall} and \texttt{mexists}) that mimic the notations
for \Coq's usual quantifiers (i.e. \texttt{forall} and
\texttt{exists}).

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Definition A \{t: Type\}(p: t -> o)(w: i) := \fforall x, p x w.
\Notation "'mforall'  x , p" := (A (\fun x => p))
  (at level 200, x ident, right associativity) : type_scope.
\Notation "'mforall' x : t , p" := (A (\fun x:t => p))
  (at level 200, x ident, right associativity, 
    format "'[' 'mforall' '/ '  x  :  t , '/ '  p ']'")
  : type_scope.

\Definition E \{t: Type\}(p: t -> o)(w: i) := exists x, p x w.
\Notation "'mexists' x , p" := (E (\fun x => p))
  (at level 200, x ident, right associativity) : type_scope.
\Notation "'mexists' x : t , p" := (E (\fun x:t => p))
  (at level 200, x ident, right associativity, 
    format "'[' 'mexists' '/ '  x  :  t , '/ '  p ']'")
  : type_scope.
\end{Verbatim}

\noindent 
The modal operators $\Diamond$ (\emph{possibly}) and $\Box$
(\emph{necessarily}) are defined accordingly to their meanings in the
possible world semantics. $\nec p$ holds at a world $w$ iff $p$ holds
in every world $w_1$ reachable from $w$. $\pos p$ holds at world $w$
iff $p$ holds in some world $w_1$ reachable from $w$.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Definition box (p: o) := \fun w => \fforall w1, (r w w1) -> (p w1).
\Definition dia (p: o) := \fun w => exists w1, (r w w1) /\bslash (p w1).
\end{Verbatim}

\noindent
A modal proposition is valid iff it holds in every possible world. 
This notion of modal validity is encoded by the following defined predicate:

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Definition V (p: o) := \fforall w, p w.
\end{Verbatim}

\noindent 
To prove a modal proposition \texttt{p} (of type \texttt{o})
within \Coq, the proposition \texttt{(V p)} (of type \texttt{Prop})
should be proved instead. To increase the transparency of the
embedding to the user, the following notation is provided, allowing
\texttt{[ p ]} to be written instead of \texttt{(V p)}.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Notation "[ p ]" := (V p).
\end{Verbatim}

\section{Tactics for Modalities}

Interactive theorem proving in \Coq is usually done with tactics,
imperative commands that reduce the theorem to be proven (i.e. the
goal) to simpler subgoals, in a bottom-up manner. The simplest tactics
can be regarded as rules of a natural deduction calculus\footnote{The
underlying proof system of \Coq (the Calculus of Inductive
Constructions \cite{Paulin}) is actually more sophisticated and
minimalistic than the calculus shown in Figure
\ref{fig:NaturalDeduction}. But the calculus shown here suffices for
the purposes of this paper.} (e.g. as those shown in Figure
\ref{fig:NaturalDeduction}).  For example: the \texttt{intro} tactic
can be used to apply the introduction rules for implication and for
the universal quantifier; the \texttt{apply} tactic corresponds to the
elimination rules for implication and for the universal quantifier;
\texttt{split} performs conjunction introduction; \texttt{exists} can
be used for existential quantifier introduction and \texttt{destruct}
for its elimination.

\newcommand{\s}{\qquad}

\begin{calculus}
{An arbitrary natural deduction calculus}
{fig:NaturalDeduction}
\begin{small}

\vspace{1em}

\s\s
\infer[\bot_E]{A}{ \bot }
\s
\infer[\imp_I]{A \imp B}{ B }
\s
\infer[\imp_I^n]{A \imp B}{ \infer*{B}{\infer[n]{A}{}} }
\s
\infer[\imp_E]{B}{A & A \imp B}

\vspace{2em}

\s
\infer[\neg\neg_E]{A}{ \neg\neg A }
\s\s
\infer[\wedge_I]{A \wedge B}{A & B}
\s\s
\infer[\wedge_{E_1}]{A}{A \wedge B}
\s\s
\infer[\wedge_{E_2}]{B}{A \wedge B}

\vspace{2em}

\s\s
\infer[\vee_E]{C}{A \vee B & \infer*{C}{\infer{A}{}} & \infer*{C}{\infer{B}{}}}
\s\s
\infer[\vee_{I_1}]{A \vee B}{A}
\s\s
\infer[\vee_{I_2}]{A \vee B}{B}

\vspace{2em}

%\begin{scriptsize}

\s\s\s
\infer[\all_I]{\all x_{\tau}. A[x]}{ A[\alpha] } 
\s\s
\infer[\all_E]{A[t]}{ \all x_{\tau}. A[x] }

\vspace{1em}

\s\s\s
\infer[\ex_I]{\ex x_{\tau}. A[x]}{ A[t] }
\s\s
\infer[\ex_E]{C}{ \ex x_{\tau}. A[x] & \infer*{C}{A[\alpha]} }

%\end{scriptsize}

\vspace{2em}

\begin{center}
$\alpha$ must respect the usual \emph{eigen-variable conditions}.
\end{center}

\end{small}

\vspace{1em}
\end{calculus}

To maximally preserve user intuition in interactive modal logic
theorem proving, the embedding via the possible world semantics should
be as transparent as possible to the user. Fortunately, the basic Coq
tactics described above automatically unfold the shallowest modal
definition in the goal.  Therefore, they can be used with modal
connectives and quantifiers just as they are used with the usual
connectives and quantifiers. The situation for the new modal
operators, on the other hand, is not as simple, unfortunately.

Since the modal operators are, in our embedding, essentially just
abbreviations for quantifiers guarded by reachability conditions, the
typical tactics for quantifiers can be used, in principle.  However,
this exposes the user to the technicalities of the embedding,
requiring him to deal with possible worlds and their reachability
explicitly. In order to obtain transparency also for the modal
operators, we have implemented a few specialized tactics using \Coq's
Ltac language.

When applied to a goal of the form \texttt{((box p) w0)}, the tactic
\texttt{box\_i} will introduce a fresh new world \texttt{w} and then
introduce the assumption that \texttt{w} is reachable from
\texttt{w0}. The new goal will be \texttt{(p w)}.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Ltac box_i := \llet w := fresh "w" \iin \llet R := fresh "R" 
               \iin (intro w at top; intro R at top).
\end{Verbatim}

\noindent 
If the hypothesis \texttt{H} is of the form \texttt{((box p)
w0)} and the goal is of the form \texttt{(q w)}, the tactic
\texttt{box\_e H H1} creates a new hypothesis \texttt{H1: (p w)}. The
tactic \texttt{box\_elim H w1 H1} is an auxiliary tactic for
\texttt{box\_e}. It creates a new hypothesis \texttt{H1: (p w1)}, for
any given world \texttt{w1}, not necessarily the goal's world
\texttt{w}. It is also responsible for automatically trying (by
\texttt{assumption}) to solve the reachability guard conditions,
releasing the user from this burden.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Ltac box_elim H w1 H1 := \match type of H \with 
      ((box ?p) ?w) => cut (p w1); 
                       [intros H1 | (apply (H w1); try assumption)] \eend.
\end{Verbatim}

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Ltac box_e H H1:= \match goal \with | [ |- (_ ?w) ] => box_elim H w H1 \eend.
\end{Verbatim}

\noindent 
If the hypothesis \texttt{H} is of the form \texttt{((dia p)
w0)}, the tactic \texttt{dia\_e H} generates a new hypothesis
\texttt{H: (p w)} for a fresh new world \texttt{w} reachable from
\texttt{w0}.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Ltac dia_e H := \llet w := fresh "w" \iin \llet R := fresh "R" \iin 
                (destruct H \as [w [R H]]; move w at top; move R at top).
\end{Verbatim}

\noindent
The tactic \texttt{dia\_i w} transforms a goal of the form 
\texttt{((dia p) w0)} into the simpler goal \texttt{(p w)} 
and automatically tries to solve the guard condition that 
\texttt{w} must be reachable from \texttt{w0}.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Ltac dia_i w := (exists w; split; [assumption | idtac]).
\end{Verbatim}

\noindent If the new modal tactics above are regarded from a natural
deduction point of view, they give rise to the inference
rules\footnote{The natural deduction calculus with the rules from
Figures \ref{fig:NaturalDeduction} and \ref{fig:NDK} is sound and
complete relatively to the calculus of Figure
\ref{fig:NaturalDeduction} extended with a necessitation rule and the
modal axiom K \cite{ToDo:UnpublishedPaperInGithub}. Therefore,
assuming that the calculus from \ref{fig:NaturalDeduction} is sound
and Henkin-complete for classical higher-order logic, the additional
modal rules in Figure \ref{fig:NDK} make it sound and Henkin-complete
for modal higher-order logic \textbf{K}.} shown in Figure
\ref{fig:NDK}. The labels that name boxes in those inference rules are
precisely the worlds that annotate goals and hypotheses in \Coq with
the modal embedding. A hypothesis of the form \texttt{(p w)}, where
\texttt{p} is a modal proposition of type \texttt{o} and \texttt{w} is
a world of type \texttt{i} indicates that \texttt{p} is an assumption
created inside a box with name \texttt{w}.

\begin{calculus}
{Rules for Modal Operators}
{fig:NDK}

\vspace{1em}

\s\s\s\s
\infer[\nec_I]{\nec A}{\omega: \fbox{\infer*{A}{}} }
\s\s\s\s
\infer[\nec_E]{w: \fbox{ \infer*{}{A} } }{\nec A}

\vspace{2em}

\s\s\s\s
\infer[\pos_I]{\pos A}{w: \fbox{\infer*{A}{}} }
\s\s\s\s\s
\infer[\pos_E]{\omega: \fbox{ \infer*{}{A} } }{\pos A}

\vspace{1em}

\begin{center}
\textbf{eigen-box condition:}\\ 
$\omega$ must be a fresh name for a box. \\
\vspace{0.5em}
\textbf{boxed assumption condition:} \\
assumptions should be discharged within the box where they are created.
\end{center}

\vspace{1em}
\end{calculus}

Finally, our implementation also provides the tactic \texttt{mv},
standing for \emph{modal validity}, which replaces a goal of the form
\texttt{[ p ]} (or equivalently \texttt{(V p)}) by a goal of the form
\texttt{(p w)} for a fresh arbitrary world \texttt{w}.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Ltac mv := \match goal \with [|- (V _)] => intro \eend.
\end{Verbatim}

\section{Convenient Modal Lemmas}

\noindent 
The following lemmas are quite convenient. The first is a
kind of modus ponens applicable to formulas under the scope of a modal
operator. The second allows pushing negations inside modalities.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Lemma mp_dia: 
  [mforall p, mforall q, (dia p) m-> (box (p m-> q)) m-> (dia q)].
\Proof. mv. 
intros p q H1 H2. dia_e H1. dia_i w0. box_e H2 H3. apply H3. exact H1. 
\Qed.

\Lemma not_dia_box_not: [mforall p, (m~ (dia p)) m-> (box (m~ p))].
\Proof. mv. 
intro p. intro H. box_i. intro H2. apply H. dia_i w0. exact H2. 
\Qed.
\end{Verbatim}

\noindent
The most well-known modal logics can be distinguished by their acceptance of the following reachability axioms. \textbf{S5} is the modal logic that accepts all three.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Axiom reflexivity: \fforall w, r w w.
\Axiom transitivity: \fforall w1 w2 w3, (r w1 w2) -> (r w2 w3) -> (r w1 w3).
\Axiom symmetry: \fforall w1 w2, (r w1 w2) -> (r w2 w1).
\end{Verbatim}

\noindent
Modal logic ``axioms'' can be derived from the reachability axioms.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Theorem K: 
  [ mforall p, mforall q, (box (p m-> q)) m-> (box p) m-> (box q) ].
\Proof. mv. 
intros p q H1 H2. box_i. box_e H1 H3. apply H3. box_e H2 H4. exact H4. 
\Qed.

\Theorem T: [ mforall p, (box p) m-> p ].
\Proof. mv. 
intro p. intro H. box_e H H1. exact H1. apply reflexivity. 
\Qed.
\end{Verbatim}

\noindent
In strong modal logics, such as S5, iterations of modal operators can be collapsed. This is a controversial principle used in G\"odel's and Scott's versions of the proof.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Theorem dia_box_to_box: [ mforall p, (dia (box p)) m-> (box p) ].
\Proof. mv. 
intros p H1. dia_e H1. box_i. box_e H1 H2. exact H2. eapply transitivity. 
  apply symmetry. exact R. 
  exact R0.
\Qed.
\end{Verbatim}

\section{G\"odel's Proof of God's Existence}
\label{sec:proof}

Attempts to prove the existence (or non-existence) of God by means of
abstract ontological arguments are an old tradition in philosophy and
theology.  
G\"{o}del's proof \cite{Goedel1970,GoedelNotes} is a modern culmination of
this tradition, following particularly the footsteps of Leibniz.
Various slightly different versions of axioms and definitions have
been considered by G\"{o}del and by several philosophers who commented
on his proof
(cf. \cite{sobel2004logic,AndersonGettings,Fitting,Adams,ContemporaryBibliography}). The formalization shown in this section aims at
being as similar as possible to Dana Scott's version of the proof \cite{ScottNotes}. The formulation and numbering of axioms, definitions and theorems is the same as in Scott's notes. Even the Coq proof scripts follow all the steps in Scott's notes. Scott's assertions are emphasized with comments. In contrast to the formalization in Isabelle \cite{AFP}, where automation via Metis \cite{Hurd03first-orderproof} and Sledgehammer \cite{Sledgehammer} using tools such as Nitpick \cite{Nitpick}, LEO-II \cite{LEO-II} and Satallax \cite{Satallax} has been successfully employed, the formalization in Coq used no automation. This was a deliberate choice, mainly because it allowed a qualitative evaluation of the convenience of the possible world semantic embedding approach for \emph{interactive} theorem proving. Moreover, in order to formalize precisely Scott's version and not another automatically found version, automation would have to be heavily limited anyway. Furthermore, the deliberate preference for simple tactics (mostly \emph{intro}, \emph{apply} and the modal tactics described in the previous section) results in proof scripts that could be read as natural deduction proofs. This hopefully makes the formalization more accessible to those who are not experts in Coq's tactics but are nevertheless interested in G\"odel's proof.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Require \Import Coq.Logic.Classical Coq.Logic.Classical_Pred_Type Modal.

\Ltac proof_by_contradiction H := apply NNPP; intro H.

\com{(* Constant predicate that distinguishes positive properties *)}
\Parameter Positive: (u -> o) -> o.

\com{(* Axiom A1:} 
\com{   either a property or its negation is positive, but not both *)}
\Axiom axiom1a : 
  [ mforall p, (Positive (\fun x: u => m~(p x))) m-> (m~ (Positive p)) ].

\Axiom axiom1b : 
  [ mforall p, (m~ (Positive p)) m-> (Positive (\fun x: u => m~ (p x))) ].

\com{(* Axiom A2: }
\com{   a property necessarily implied by a positive property is positive *)}
\Axiom axiom2: [ mforall p, mforall q, 
  Positive p m/\bslash (box (mforall x, (p x) m-> (q x) )) m-> Positive q ].

\com{(* Theorem T1: positive properties are possibly exemplified *)}
\Theorem theorem1: [ mforall p, (Positive p) m-> dia (mexists x, p x) ].
\Proof. mv.
intro p. intro H1. proof_by_contradiction H2. apply not_dia_box_not in H2.
assert (H3: ((box (mforall x, m~ (p x))) w)). \com{(* Scott *)}
  box_i. intro x. assert (H4: ((m~ (mexists x : u, p x)) w0)).
    box_e H2 G2. exact G2.
    clear H2 R H1 w. intro H5. apply H4. exists x. exact H5.
  assert (H6: ((box (mforall x, (p x) m-> m~ (x m= x))) w)). \com{(* Scott *)}   
    box_i. intro x. intros H7 H8. box_elim H3 w0 G3. eapply G3. exact H7.
    assert (H9: ((Positive (fun x => m~ (x m= x))) w)). \com{(* Scott *)}
      apply (axiom2 w p (fun x => m~ (x m= x))). split.
        exact H1.
        exact H6.
      assert (H10: ((box (mforall x, (p x) m-> (x m= x))) w)). \com{(* Scott *)}
        box_i. intros x H11. reflexivity.
        assert (H11 : ((Positive (fun x => (x m= x))) w)). \com{(* Scott *)}
          apply (axiom2 w p (fun x => x m= x )). split.
            exact H1.
            exact H10.
          apply axiom1a in H9. contradiction.
\Qed.

\com{(* Definition D1: }
\com{   God: a God-like being possesses all positive properties *)}
\Definition G(x: u) := mforall p, (Positive p) m-> (p x).

\com{(* Axiom A3: the property of being God-like is positive *)}
\Axiom axiom3: [ Positive G ].

\com{(* Corollary C1: possibly, God exists *)}
\Theorem corollary1: [ dia (mexists x, G x) ]. 
\Proof. mv. apply theorem1. apply axiom3. \Qed.

\com{(* Axiom A4: positive properties are necessarily positive *)}
\Axiom axiom4: [ mforall p, (Positive p) m-> box (Positive p) ].

\com{(* Definition D2: }
\com{   essence: an essence of an individual is a property possessed by it} 
\com{   and necessarily implying any of its properties *)}
\Definition Essence(p: u -> o)(x: u) := 
    (p x) m/\bslash mforall q, ((q x) m-> box (mforall y, (p y) m-> (q y))).
\Notation "p 'ess' x" := (Essence p x) (at level 69).

\com{(* Theorem T2: being God-like is an essence of any God-like being *)}
\Theorem theorem2: [ mforall x, (G x) m-> (G ess x) ].
\Proof. mv. intro g. intro H1. unfold Essence. split.
  exact H1.
  intro q. intro H2. assert (H3: ((Positive q) w)).
    proof_by_contradiction H4. unfold G in H1. apply axiom1b in H4. 
    apply H1 in H4. contradiction.

    cut (box (Positive q) w). \com{(* Scott *)}
      apply K. box_i. intro H5. intro y. intro H6. 
      unfold G in H6. apply (H6 q). exact H5.

      apply axiom4. exact H3.
\Qed.

\com{(* Definition D3: }
\com{   necessary existence: necessary existence of an individual }
\com{   is the necessary exemplification of all its essences *)}
\Definition NE(x: u) := mforall p, (p ess x) m-> box (mexists y, (p y)).

\com{(* Axiom A5: necessary existence is a positive property *)}
\Axiom axiom5: [ Positive NE ].

\Lemma lemma1: [ (mexists z, (G z)) m-> box (mexists x, (G x)) ].
\Proof. mv. 
intro H1. destruct H1 as [g H2]. cut ((G ess g) w). \com{(* Scott *)}
  assert (H3: (NE g w)).       \com{(* Scott *)}
    unfold G in H2. apply (H2 NE). apply axiom5.
    unfold NE in H3. apply H3.
  apply theorem2. exact H2.
\Qed.

\Lemma lemma2: [ dia (mexists z, (G z)) m-> box (mexists x, (G x)) ].
\Proof. mv. 
intro H. cut (dia (box (mexists x, G x)) w).  \com{(* Scott *)}
  apply dia_box_to_box.
  apply (mp_dia w (mexists z, G z)).
    exact H.
    box_i. apply lemma1.
\Qed.

\com{(* Theorem T3: necessarily, a God exists *)}
\Theorem theorem3: [ box (mexists x, (G x)) ].
\Proof. mv. apply lemma2. apply corollary1. \Qed.

\com{(* Corollary C2: There exists a god *)}
\Theorem corollary2: [ mexists x, (G x) ].
\Proof. mv. apply T. apply theorem3. \Qed.
\end{Verbatim}

\section{Conclusions}
\label{sec:conclusions}

The successful formalization of Scott's version of G\"odel's proof of
God's existence indicates that the \emph{embedding} of modal logics
into higher-order logics via the \emph{possible world semantics} is a
viable approach for interactive theorem proving within modal logics.
The minimalistic implementation of the embedding (Section
\ref{sec:modal}) takes special care to hide the underlying possible
world machinery from the user.  An inspection of the proof scripts in
Section \ref{sec:proof} shows that this goal has been achieved. The
user does not have to explicitly bother about worlds and their mutual
reachability; the provided tactics for modalities do the job for him/her.
Moreover, for subgoals that do not involve modalities, the user has all the
usual interactive tactics at his/her disposal.
%The implementation of hints to allow automation tactics to fully cope with the embedding remains for future work. Nevertheless, at this stage a tactic such as \texttt{firstorder} is able to prove all lemmas and theorems in Section \ref{sec:modal} fully automatically.

HOL Light, Isabelle

The theological implications of the verified correctness of this proof
certainly depend on a critical discussion of the underlying concepts,
definitions and axioms, which is beyond the scope of this paper. 
%What
%can be said though is that our work presents further evidence that
%theism is at least not (obviously) irrational. 
Clearly, the application of theorem proving technology to the domain of
theoretical philosophy can --- as already pictured by Leibniz ---
be very fruitful for both areas. In fact, new results of
philosophical interest have been obtained by our
work~\cite{arXiv}. Moreover a first, basic infrastructure for
interactive and automated reasoning in higher-order modal logics has
been created and will be further improved in the future.

\paragraph{Acknowledgements:} we thanks Cedric Auger and Laurent Th\'ery, for their answers to our questions about Ltac in the Coq-Club mailing-list.

%\footnotesize
\bibliographystyle{plain}
%\bibliography{Bibliography}
\begin{thebibliography}{10}

\bibitem{Adams}
R.M.~Adams.
\newblock Introductory note to *1970.
\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
  Letters}}. Oxford University Press, 1995.

\bibitem{AndersonGettings}
A.C.~Anderson and M.~Gettings.
\newblock G\"odel ontological proof revisited.
\newblock In {\em {G\"odel'96: Logical Foundations of Mathematics, Computer
  Science, and Physics: Lecture Notes in Logic 6}}. {Springer}, 1996.

\bibitem{B9}
C.~Benzm{\"u}ller and L.C.~Paulson.
\newblock Exploring properties of normal multimodal logics in simple type
  theory with {LEO-II}.
\newblock In {\em {Festschrift in Honor of {Peter B. Andrews} on His 70th
  Birthday}}, pages 386--406. College Publications.

\bibitem{J23}
C.~Benzm{\"u}ller and L.C.~Paulson.
\newblock Quantified multimodal logics in simple type theory.
\newblock {\em Logica Universalis (Special Issue on Multimodal Logics)},
  7(1):7--20, 2013.

\bibitem{LEO-II}
C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke.
\newblock {LEO-II} - a cooperative automatic theorem prover for higher-order
  logic.
\newblock In {\em Proc. of IJCAR 2008}, volume 5195 of {\em LNAI}, pages
  162--170. Springer, 2008.

\bibitem{arXiv}
C.~Benzm{\"u}ller and B.~Woltzenlogel Paleo,
\newblock Formalization, Mechanization and Automation of G{\"o}del's Proof of God's Existence.
\newblock {\em {arXiv}}:1308.4526, 2013.

\bibitem{AFP}
C.~Benzm{\"u}ller and B.~Woltzenlogel Paleo,
\newblock G{\"o}del's God in Isabelle/HOL. 
\newblock {\em Archive of Formal Proofs}, 2013.

\bibitem{Coq}
Y.~Bertot and P.~Casteran.
\newblock {\em {Interactive Theorem Proving and Program Development}}.
\newblock Springer, 2004.

\bibitem{ModalLogic}
P.~Blackburn, J.v.~Benthem, and F.~Wolter (eds.), \newblock {\em {Handbook of Modal Logic}}. Elsevier, 2006.

\bibitem{Sledgehammer}
J.C.~Blanchette, S.~B\"ohme, and L.C.~Paulson.
\newblock Extending {Sledgehammer} with {SMT} solvers.
\newblock {\em Journal of Automated Reasoning}, 51(1):109--128, 2013.

\bibitem{Nitpick}
J.C.~Blanchette and T.~Nipkow.
\newblock Nitpick: A counterexample generator for higher-order logic based on a
  relational model finder.
\newblock In {\em Proc. of ITP 2010}, no.~6172 in LNCS, pages 131--146.
  Springer, 2010.

\bibitem{Zalta}
P.E.~Oppenheimera and E.N.~Zalta.
\newblock A Computationally-Discovered Simplification of the Ontological Argument.
\newblock {\em Australasian Journal of Philosophy}, 89(2):333-349, 2011.

\bibitem{Satallax}
C.E.~Brown.
\newblock Satallax: An automated higher-order prover.
\newblock In {\em Proc. of IJCAR 2012}, number 7364 in LNAI, pages 111 -- 117.
  Springer, 2012.

\bibitem{ContemporaryBibliography}
R.~Corazzon.
\newblock Contemporary bibligraphy on the ontological proof
  (\url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}).

\bibitem{Fitting}
M.~Fitting.
\newblock {\em {Types, Tableaux and G\"odel's God}}.
\newblock Kluver Academic Press, 2002.

\bibitem{Labels}
D.M.~Gabbay. {\em Labelled Deductive Systems}. Clarendon Press, 1996.

\bibitem{Goedel1970}
K.~G\"odel.
\newblock Ontological proof.
\newblock In {\em {Kurt G\"odel: Collected Works Vol. 3: Unpublished Essays and
  Letters}}. Oxford University Press, 1970.

\bibitem{GoedelNotes}
K.~G\"odel.
\newblock {\em Appendix A. Notes in Kurt G\"odel's Hand}, pages 144--145.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{Hazen}
A.P.~Hazen.
\newblock On g\"odel's ontological proof.
\newblock {\em Australasian Journal of Philosophy}, 76:361--377, 1998.

\bibitem{Hurd03first-orderproof}
J.~Hurd.
\newblock First-order proof tactics in higher-order logic theorem provers.
\newblock In {\em Design and Application of Strategies/Tactics in Higher Order
  Logics, NASA Tech. Rep. NASA/CP-2003-212448}, 2003.

\bibitem{Isabelle}
T.~Nipkow, L.C. Paulson, and M.~Wenzel.
\newblock {\em {Isabelle/HOL: A Proof Assistant for Higher-Order Logic}}.
\newblock Number 2283 in LNCS. Springer, 2002.

\bibitem{Rushby}
J.~Rushby.
\newblock The Ontological Argument in PVS. 
\newblock In {\em Proc.~of CAV Workshop ``Fun With Formal Methods''}, St. Petersburg, Russia, 2013.

\bibitem{ScottNotes}
D.~Scott.
\newblock {\em Appendix B. Notes in Dana Scott's Hand}, pages 145--146.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{sobel2004logic}
J.H.~Sobel.
\newblock {\em Logic and Theism: Arguments for and Against Beliefs in God}.
\newblock Cambridge U. Press, 2004.

\bibitem{J22}
G.~Sutcliffe and C.~Benzm{\"u}ller.
\newblock Automated reasoning in higher-order logic using the {TPTP THF}
  infrastructure.
\newblock {\em Journal of Formalized Reasoning}, 3(1):1--27, 2010.

\bibitem{FormalTheologyRepository}
B.~Woltzenlogel Paleo and C.~Benzm\"uller.
\newblock Formal theology repository
  (\url{https://github.com/FormalTheology/GoedelGod}).
\end{thebibliography}

\end{document}