### 2015-07-16 Talks/2015/CSR/talk-part-1.tex
Add CSR Talk

% \begin{transitionframe}{Images/Transitions/CartoucheThutmosisReverse(WaltersArtMuseum)(PublicDomain)}{black} \centering
% \textbf{Embedding Modal Logic into Higher-Order Logic}
% \end{transitionframe}

\begin{frame}{Quick Introduction to Modal Logics}{Modal Operators} \large
\centering

$$\nec P$$
P is necessary, P is obligatory, P is known, \\ 
P is believed, always P \ldots

\bigskip
\pause

$$\pos P$$
P is possible, P is permissible, P is epistemically possible, \\ 
P is doxastically possible, eventually P \ldots

\end{frame}

\begin{frame}{Quick Introduction to Modal Logics}{Kripke Semantics - Possible Worlds} \large

\includegraphics[width=0.8\textwidth]{Images/Kripke}

\end{frame}

\begin{frame}{Quick Introduction to Higher-Order Logics}{Higher-Order Quantification} \large

\begin{Huge}
$$
\forall x. (G(x) \equiv \alert{\forall \varphi}. P(\varphi) \imp \alert{\varphi}(x))
$$
\end{Huge}

\end{frame}

\begin{frame}{Quick Introduction to Higher-Order Logics}{Lambda Calculus: Types and Beta-Reduction} \large

\begin{Huge}
$$
(\lambda \varphi_{\alert{\iota \imp o}}. (P ~ \varphi) ) ~ G  \qquad  \leadsto_{\beta} \qquad (P ~ G)
$$
\end{Huge}
\end{frame}

\begin{frame}{Motivation, Challenge, Goals, Approach} \large

\visible<+->{
\hskip-1em Motivation : Applications (Ontologies, Paraconsistency, \emph{Philosophy}, \ldots) \\[1em]
}

\visible<+->{
\hskip-1em Challenge : No provers for \emph{Higher-order
  Quantified Modal
  Logics\/} (\textcolor{red}{QML}) \\[1em]
}

\visible<+->{
\hskip-1em Our goals : \\
  \qquad efficient automated provers for \textcolor{red}{QML} \\
  \qquad user-friendly interactive provers for \textcolor{red}{QML} \\[1em]
}

\visible<+->{
\hskip-1em Approach : Embed \textcolor{red}{QML} in \emph{Higher-Order Classical
  Logic\/} (\textcolor{blue}{HOL}) \\
\, \hfill Then use existing \textcolor{blue}{HOL} theorem provers for reasoning in \textcolor{red}{QML} \\
  \qquad\qquad\qquad interactive: \hfill Isabelle, HOL4, Hol Light, Coq, PVS, \ldots \\
  \qquad\qquad\qquad automated: \hfill TPS, LEO-II, Satallax, Nitpick, \ldots \\
}
\end{frame}

\begin{frame}{Embedding QML in HOL}\large

\hskip-1em\textcolor{red}{QML} \hfill
$\begin{array}{lll}\textcolor{red}{\varphi,\psi} & ::= &
  \textcolor{red}{\ldots}  \mid \textcolor{red}{\neg
    \varphi} \mid \textcolor{red}{\varphi \wedge \psi} \mid
  \textcolor{red}{\varphi \imp \psi}  \mid \textcolor{red}{\Box
    \varphi} \mid \textcolor{red}{\Diamond \varphi}  \mid
  \textcolor{red}{\forall {x}\, \varphi} \mid
  \textcolor{red}{\exists {x}\, \varphi} 
\mid \textcolor{red}{\forall {P}\, \varphi} \end{array}$ \\[1em]

\hskip-1em\textcolor{blue}{HOL}\hfill 
$\begin{array}{lll}
\textcolor{blue}{s,t} & ::= & \textcolor{blue}{C}  \mid
\textcolor{blue}{x \mid \lambda{x} s} \mid \textcolor{blue}{s\, t}
\mid \textcolor{blue}{\neg s} \mid \textcolor{blue}{s \vee t} \mid
\textcolor{blue}{\forall {x}\, t} 
\end{array}$ \\[1em]

\pause

\hskip-1em\textcolor{red}{QML} in \textcolor{blue}{HOL}: \quad \textcolor{red}{QML}
formulas $\textcolor{red}{\varphi}$ are mapped to
\textcolor{blue}{HOL} predicates $\textcolor{red}{\varphi_{\worldtype\typearrow o}}$

\begin{center}
\fcolorbox{blue}{white}{
$\begin{array}{lcl} 
    \textcolor{red}{\mnot} & = & \textcolor{blue}{
      \lambda{\varphi_{\worldtype\typearrow o}}\lambda{w_\worldtype}\neg \varphi w} \\ 
    \textcolor{red}{\mand} & = & \textcolor{blue}{ 
      \lambda{\varphi_{\worldtype\typearrow o}}
      \lambda{\psi_{\worldtype\typearrow o}} \lambda{w_\worldtype}
      (\varphi w \wedge \psi w)} \\ 
    \textcolor{red}{\imp} & = & \textcolor{blue}{ 
      \lambda{\varphi_{\worldtype\typearrow o}}
      \lambda{\psi_{\worldtype\typearrow o}} \lambda{w_\worldtype}
      (\neg \varphi w \vee \psi w)} \\ 
    \textcolor{red}{\forall} & = & \textcolor{blue}{ 
      \lambda{h_{\gamma\typearrow(\worldtype \typearrow o)}}
      \lambda{w_\worldtype} \forall {d_\gamma} \, h d w} \\
    \textcolor{red}{\exists} & = & \textcolor{blue}{ 
      \lambda{h_{\gamma\typearrow(\worldtype \typearrow o)}}
      \lambda{w_\worldtype} \exists {d_\gamma} \, h d w} \\
    \\
    \textcolor{red}{\Box} & = & \textcolor{blue}{ 
      \lambda{\varphi_{\worldtype\typearrow o}} \lambda{w_\worldtype}
      \forall {u_\worldtype}\, (\neg r w u \vee
      \varphi u)} \\ 
    \textcolor{red}{\Diamond} & = & \textcolor{blue}{ 
      \lambda{\varphi_{\worldtype\typearrow o}} \lambda{w_\worldtype}
      \exists {u_\worldtype}\, (r w u \wedge
      \varphi u)} \\ 
    \\
    \text{\textcolor{brown}{valid}} & = & \textcolor{blue}{
        \lambda{\varphi_{\worldtype\typearrow o}} \all{w_\worldtype}
        \varphi w}
\end{array}$
}  \quad \textcolor{blue}{Ax} 
\vskip1em
\end{center}

\pause

\quad The equations in \textcolor{blue}{Ax} are given as axioms to the \textcolor{blue}{HOL} provers. \\

\end{frame}

\begin{frame}{Embedding QML in HOL} \large

\hskip-1em Example \\[.5em]

 \textcolor{red}{QML} formula  \hfill \textcolor{red}{$\Diamond \exists x G(x)$}

\pause

 \textcolor{red}{QML} formula in \textcolor{blue}{HOL}  \hfill $\text{\textcolor{brown}{valid}}\, \textcolor{red}{(\Diamond \exists x G(x))_{\worldtype\typearrow o}}$

\pause

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype\textcolor{red}{(\Diamond \exists x
    G(x))_{\worldtype\typearrow o}}\, w}$

\pause

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype \exists {u_\worldtype} (r w u \wedge
      \textcolor{red}{(\exists x G(x))_{\worldtype\typearrow o}} u)}$

% expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
%   w_\worldtype \exists {u_\worldtype} (r w u \wedge
%       \exists x \textcolor{red}{G(x)_{\worldtype\typearrow o}} u)}$

\pause

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype \exists {u_\worldtype} (r w u \wedge
      \exists x G x u)}$ \\[1em]

\pause

\visible<6->{
\vfill
\begin{block}{What do we do?}
\vskip.5em
In order to prove that $\textcolor{red}{\varphi}$ is valid in \textcolor{red}{QML}, \\
we instead prove that 
$\text{\textcolor{brown}{valid}}\,
\textcolor{red}{\varphi_{\worldtype\typearrow o}}$ can be derived
from \textcolor{blue}{Ax} in \textcolor{blue}{HOL}. \\[1em]

This can be done with interactive or automated \textcolor{blue}{HOL} theorem provers.
\end{block}
}

\end{frame}

\begin{frame}{Embedding QML in HOL} \large

\hskip-1em Example \\[.5em]

 \textcolor{red}{QML} formula  \hfill \textcolor{red}{$\Diamond \exists x G(x)$}

 \textcolor{red}{QML} formula in \textcolor{blue}{HOL}  \hfill $\text{\textcolor{brown}{valid}}\, \textcolor{red}{(\Diamond \exists x G(x))_{\worldtype\typearrow o}}$

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype\textcolor{red}{(\Diamond \exists x
    G(x))_{\worldtype\typearrow o}}\, w}$

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype \exists {u_\worldtype} (r w u \wedge
      \textcolor{red}{(\exists x G(x))_{\worldtype\typearrow o}} u)}$

% expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
%   w_\worldtype \exists {u_\worldtype} (r w u \wedge
%       \exists x \textcolor{red}{G(x)_{\worldtype\typearrow o}} u)}$

expansion, $\beta\eta$-conversion \hfill $\textcolor{blue}{\forall
  w_\worldtype \exists {u_\worldtype} (r w u \wedge
      \exists x G x u)}$ \\[1em]

\vfill
\begin{block}{Does this actually work in practice?}
\vskip.5em
\centering
Is it efficient ??

\bigskip

Is it user-friendly ??

\bigskip

\end{block}

\end{frame}