### 2015-07-16 Talks/2015/CSR/talk-part-4.tex

% \begin{transitionframe}{Images/Transitions/TrifidNebula(Nasa)(PublicDomain).jpg}{black}
% \textbf{Conclusions and Future Work}
% \end{transitionframe}

\begin{frame}{Conclusions} \large

\begin{itemize}
\item Contributions: \\[.5em]
\begin{itemize}
\item Efficient automated reasoning for QML \\[.3em]
\item \alert{User-friendly interactive reasoning with QML in Coq} \\[.3em]
\item \alert{A new natural deduction calculus for higher-order modal logics} \\[.3em]
\item Non-trivial new benchmark problems for HOL provers \\[.3em]
\item Verification of existing results about G\"odel's proof (e.g. Modal Collapse) \\[.3em]
\item New results about G\"odel's proof (e.g. Inconsistency) \\[.3em]
\item Resolution of philosophical controversies (Leibniz's dream) \\[.3em]
\item Computer-Assisted Theoretical Philosophy \\[.3em]
 \begin{itemize}
  \item Ed Zalta's \emph{Computational Metaphysics} project at Stanford University \\[.2em]
  \item John Rushby's formalization of Anselm's proof using PVS
  \end{itemize}
\end{itemize}

\end{itemize}
\end{frame}